\subsection{Direct Proof and Counterexample}
\hrulefill

\paragraph*{Definitions}
Let P(n) be the predicate "n is an even number".
\begin{align*}
    \forall n \in \mathbb{Z}, P(n) \iff \exists k \in \mathbb{Z} \ni n = 2k.\\
    \forall n \in \mathbb{Z}, \neg P(n) \iff \exists k \in \mathbb{Z} \ni n = 2k + 1.
\end{align*}

\paragraph*{Example}
Is -301 even or odd?
\begin{align*}
    -301 &= 2k + 1 \text{ for } k=-151\\
\end{align*}

\paragraph*{Example}
If $a, b \in \mathbb{Z}$, is $6a^2b$ even?
\begin{align*}
    &\exists a, b \in \mathbb{Z} \ni 6a^2b = 2(k) + 1 \\
    &6a^2b = 2(3a^2b) \text{ for } k=3a^2b\\
    \therefore \quad &6a^2b \text{ is even.}
\end{align*}

\paragraph*{Prime and Composite Number Definition}
Let P(n) be the predicate "n is a prime number".
\begin{align*}
    \forall n \in \mathbb{Z}_{>1}, P(n) &\iff \forall r, s \in \mathbb{Z}_{>1}, n=rs \implies r=n \lor s=n\\
    \forall n \in \mathbb{Z}_{>1}, \neg P(n) &\iff \exists r, s \in \mathbb{Z}_{>1} \ni n=rs \land 1 < r < n \land 1 < s < n
\end{align*}

\subsubsection*{Constructive Proof of Existential Statement}
\begin{equation*}
    \exists x in D \ni Q(x)
\end{equation*}
\begin{itemize}
    \item Find an x in D that makes !(x) true.
    \item Give a set of directions for finding such an x in D
\end{itemize}

\paragraph*{Example}
Prove there is and even integer n such that n can be written in two ways as a sum of two prime numbers.
\begin{align*}
    \text{Let } n &= 10,\\
    10 &= 3 + 7\\
    10 &= 5+5\\
    \therefore \quad &\text{the statement is true.}
\end{align*}

\subsubsection*{Disproving Universal Statement by Counterexample}
\begin{equation*}
    \forall x in D, P(x) \implies Q(x)
\end{equation*}

\begin{itemize}
    \item Find an x in D that makes P(x) true, but Q(x) false.
\end{itemize}

% \paragraph*{Example}
% Disprove the statement $\forall n, k \in \mathbb{Z}, n = 2k \implies \frac{n-1}{2} = 2k + 1$
% \begin{align*}
%     Let n = 1\\
%     1 = 2(0) + 1, \therefore n is odd.
%     \frac{1-1}{2}
% \end{align*}

\subsubsection*{Method of Exhaustion of Proving Universal Statement}
\begin{equation*}
    \forall x in D, P(x) \implies Q(x)
\end{equation*}
\begin{itemize}
    \item Check all x in D to make sure that when P(x) is true, Q(x) is false.
\end{itemize}

% \paragraph*{Example}
% Prove the statement "$\forall n \in 4 \leq n \leq 10, n = a + b$ where a and b are prime numbers.
% \begin{align*}
%     n &= 4, \quad 4 &= 2+2
% \end{align*}

\subsubsection*{Direct Proof of Universal Statement}
\begin{equation*}
    \forall x \in D, P(x) \implies Q(x)
\end{equation*}
\begin{itemize}
    \item Suppose x is an arbitrary element in D for which the hypothesis P(x) is true.
    \item Using definitions or previously established results and rules to conclude Q(x) is true.
\end{itemize}

\paragraph*{Example}
Prove the statement "the sum of any two even integers is even."
\begin{align*}
    &\text{Suppose a and b are two even integers}\\
    \therefore \quad & a = 2k, \exists k_1 \in \mathbb{Z}\\
    \therefore \quad & b = 2k, \exists k_2 \in \mathbb{Z}\\
    \therefore \quad & a + b = 2k_1 + 2k_2\\
    \therefore \quad & a + b = 2(k_1 + k_2)\\
    \therefore \quad & a + b \text{ is even}
\end{align*}